Lab 13 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers

### 2023-04-18 Lab 14  RDD/Lab13_RDD.Rmd
Add files via upload

---
title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output: html_document
bibliography: rdd.bib
---

@@

#Running iv model using ivreg from the aer package
reg_iv1 <- ivreg(lwage~ educ+age+exper+expersq| 
                   .-educ+fatheduc+motheduc, data=women)

@@

#Diagnostics: 
#Conducting an F-test
#Running first_stage model
first_stage <- lm(educ~fatheduc+motheduc,data=women) #regress X on Z 
#Running null model
null <- lm(educ~1,data=women)
relevance_test <- waldtest(first_stage, null)
relevance_test

@@

For the exercise we are going to replicate the paper by @lmb2004 : \textit{Do Voters Affect or Elect Policies? Evidence from the U.S. House} in the QJE (referred to as LMB below). You don't have to read the paper, but essentially in their paper they are trying to test which model of public policy is more plausible: Do elections push elected officials somewhere to the middle (i.e. following the Downsian Model of voting), such that the resultant policy outcomes are all clustered around the median voter? Or do elections push the politician to a specific direction? If politicians cannot make credible promises to moderate their policies, then politicians would "change their minds" after an election, to follow the general ideological direction expressed by the electorate in the election.

@@

```{r}
# Load Data
enricoall4 <- read_dta("enricoall4.dta")

@@

# Plot RDD

@@

```{r}
# Fit RDD Model specifying clustered standard errors for DW-Nom as DV

# Plot RDD

@@

Yes, we have successfully replicated their results found in the paper.

## McCrary Test

```{r}
#McCrary Test

```
The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable. The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win). You can also see this above with the lack of observations around the cutpoint. This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable. Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!

# Bibliography

>>>>

---
title: "Lab 13 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output: html_document
---

@@

#Running basic model
# models wage as a function of a constant and educ
reg_ex <- lm(lwage~educ,data=women)
summary(reg_ex)

#This indicates that every additional year of education increases the wage by almost 11% 
#The issue with this sort of model is that education is most likely to be correlated with individual characteristics that are important for the person's wage, but not modelled (and hence captured by the error term).

#Enter IV!

#We'll be using parents’ education (fatheduc and motheduc) as instruments for education.
#Running iv model using ivreg() from the aer package

#Before the vertical line we can see the model that is to be estimated, lwage~educ+age+exper+expersq
#After the vertical line, we first see the instrumental variables used to instrument educ, fatheduc+motheduc; this is followed by all the explanatory variables that are considered exogenous, (age+exper+expersq), rather than re-writing them we can tell R which variables to remove from the instrument set (education).

reg_iv1 <- ivreg(lwage~ educ+age+exper+expersq| 
                   .-educ+fatheduc+motheduc, data=women)

@@

#Diagnostics: 
#Conducting an F-test
#Running first_stage model
first_stage <- lm(educ~age+exper+expersq+fatheduc+motheduc,data=women) #regress X on Z 
#Running null model
null <- lm(educ~1,data=women)
relevance_test <- waldtest(first_stage, null)
relevance_test

@@

For the exercise we are going to replicate the paper by David S. Lee, Enrico Moretti, Matthew J. Butler : \textit{Do Voters Affect or Elect Policies? Evidence from the U.S. House} in the QJE (referred to as LMB below). You don't have to read the paper, but essentially in their paper they are trying to test which model of public policy is more plausible: Do elections push elected officials somewhere to the middle (i.e. following the Downsian Model of voting), such that the resultant policy outcomes are all clustered around the median voter? Or do elections push the politician to a specific direction? If politicians cannot make credible promises to moderate their policies, then politicians would "change their minds" after an election, to follow the general ideological direction expressed by the electorate in the election.

@@

```{r, eval=FALSE}
# Load Data
enricoall4 <- read_dta("enricoall4.dta")

@@

# Plot RDD
lmb.fit <- rdrobust(enricoall4$realada,
            _________)
rdplot(____,
       ______,
       x.label="Democrat Vote Share at time t",
       y.label="ADA Score at time t+1",
       c=0.5)

@@

```{r, eval=FALSE}
# Fit RDD Model specifying clustered standard errors for DW-Nom as DV

lmb.fit2 <- rdrobust(enricoall4$dwnom1,
                     ________)

# Plot RDD
rdplot(_____,
       ______,
       x.label="Democrat Vote Share at time t",
       y.label="DW-Nominate Score at time t",
         c=0.5)

@@

```{r,eval=FALSE}
#McCrary Test

DCdensity(______)
```

### 2023-04-18 Lab 14  RDD/Lab13_rdd_Answers.Rmd
Add files via upload

---
title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output:
  pdf_document: default
  html_document: default
---

@@

```{r}

#Reading in the data
women<- read.csv("http://eclr.humanities.manchester.ac.uk/images/5/5f/Mroz.csv")

@@

#Based on the output, I find that the instruments are both relevant (confirmed by F-test) and valid (confirmed by Sargan test) to education, and that higher levels of education slightly increases a women's wage (significant at the 0.1 level).

```
# Exercise: Do Voters change the mind of their representatives? Evidence from the U.S. House.

For the exercise we are going to replicate the paper by @lmb2004 : \textit{Do Voters Affect or Elect Policies? Evidence from the U.S. House} in the QJE (referred to as LMB below). You don't have to read the paper, but essentially in their paper they are trying to test which model of public policy is more plausible: Do elections push elected officials somewhere to the middle (i.e. following the Downsian Model of voting), such that the resultant policy outcomes are all clustered around the median voter? Or do elections push the politician to a specific direction? If politicians cannot make credible promises to moderate their policies, then politicians would "change their minds" after an election, to follow the general ideological direction expressed by the electorate in the election.

To test this proposition, LMB fitted two simple regression discontinuity models. They used the ADA Score and the DW-Nominate Scores (figures I and VI respectively in the paper), both are quantitative measures of the ideological position of representatives. They argue that the running variable to the Democratic Vote Share at time $t$ - So if democrats receive more than half the share of votes, they would win in that district and a democratic representative would be elected. If elections can coerce the politician to change their policy predispositions, then at time $t+1$ we should observe a movement in the ADA Score and DW-Nominate Score (Increase in the former - converging to 100; and decrease in the latter, converging to -1).

@@

```{r}
#McCrary Test
DCdensity(enricoall4$lagdemvoteshare, 0.5)
```
The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable. The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win). You can also see this above with the lack of observations around the cutpoint. This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable. Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!

>>>>

---
title: "Lab 13 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output:
  html_document: default
  pdf_document: default
---

@@

```{r}
#Reading in the data
women<- read.csv("http://eclr.humanities.manchester.ac.uk/images/5/5f/Mroz.csv")

@@

#Based on the output, I find that the instruments are both relevant (confirmed by F-test) and valid (confirmed by Sargan test) to education, and that higher levels of education slightly increases a women's wage (significant at the 0.1 level). 
```
# Exercise: Do Voters change the mind of their representatives? Evidence from the U.S. House.

For the exercise we are going to replicate the paper by David S. Lee, Enrico Moretti, Matthew J. Butler : \textit{Do Voters Affect or Elect Policies? Evidence from the U.S. House} in the QJE (referred to as LMB below). You don't have to read the paper, but essentially in their paper they are trying to test which model of public policy is more plausible: Do elections push elected officials somewhere to the middle (i.e. following the Downsian Model of voting), such that the resultant policy outcomes are all clustered around the median voter? Or do elections push the politician to a specific direction? If politicians cannot make credible promises to moderate their policies, then politicians would "change their minds" after an election, to follow the general ideological direction expressed by the electorate in the election.

To test this proposition, LMB fitted two simple regression discontinuity models. They used the ADA Score and the DW-Nominate Scores (figures I and VI respectively in the paper), both are quantitative measures of the ideological position of representatives.

They argue that the running variable to the Democratic Vote Share at time $t$ - So if democrats receive more than half the share of votes, they would win in that district and a democratic representative would be elected. If elections can coerce the politician to change their policy predispositions, then at time $t+1$ we should observe a movement in the ADA Score and DW-Nominate Score (Increase in the former - converging to 100; and decrease in the latter, converging to -1).

@@

```{r}
#McCrary Test
DCdensity(enricoall4$lagdemvoteshare, 0.5)
```
The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable. The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win).

You can also see this above with the lack of observations around the cutpoint. This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable. Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!
